### 2023-12-12 ReplicationPackage/tables/Table4.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{6}{c}}
\hline\hline
                &\multicolumn{1}{c}{(1)}&\multicolumn{1}{c}{(2)}&\multicolumn{1}{c}{(3)}&\multicolumn{1}{c}{(4)}&\multicolumn{1}{c}{(5)}&\multicolumn{1}{c}{(6)}\\
                &\multicolumn{1}{c}{1982-2001}&\multicolumn{1}{c}{2003-2017}&\multicolumn{1}{c}{1982-2017}&\multicolumn{1}{c}{1982-2017}&\multicolumn{1}{c}{1982-2017}&\multicolumn{1}{c}{High MTR}\\
\hline
Age             &  0.00186\sym{***}& -0.00134\sym{***}&  0.00186\sym{***}& 0.000641         & -0.00453\sym{***}&0.0000502         \\
                &(0.000391)         &(0.000358)         &(0.000391)         &(0.000470)         &(0.00117)         &(0.000258)         \\
[1em]
Age X post-2001 &                  &                  & -0.00320\sym{***}& -0.00235\sym{***}& -0.00213\sym{*}  &-0.000628         \\
                &                  &                  &(0.000530)         &(0.000571)         &(0.00113)         &(0.000552)         \\
\hline
Observations    &     7743         &     5689         &    13432         &    13432         &    13432         &    12910         \\
Year Fixed Effects          &      Yes         &      Yes         &      Yes         &      Yes         &      Yes         &      Yes         \\
State Fixed Effects         &       No         &       No         &       No         &      Yes         &      Yes         &      Yes         \\
Individual Fixed Effects         &       No         &       No         &       No         &       No         &      Yes         &       No         \\
\hline \hline
\multicolumn{5}{l}{\footnotesize Driscoll-Kraay standard errors (in parentheses).} \\
\multicolumn{5}{l}{\footnotesize \sym{*} \(p<0.10\), \sym{**} \(p<0.05\), \sym{***} \(p<0.01\)}  \end{tabular} }