### 2023-12-12 ReplicationPackage/tables/Table7_a.tex

& & \multicolumn{2}{c}{\underline{Baseline Elasticity}}& \multicolumn{2}{c}{\underline{Lower Elasticity}}\\ &Personal Income & &EPV Net Revenues& &EPV Net Revenues\\State& Tax Rate&Cost/Benefit&from Adopting (mil)&Cost/Benefit&from Adopting (mil)\\
\midrule
Connecticut&6.99&0.74&3,862&0.32&11,237 \\
Delaware&6.80&0.72&417&0.31&1,143 \\
Hawaii&11.20&1.18&-472&0.52&1,446 \\
Illinois&4.35&0.46&16,035&0.20&26,856 \\
Maine&7.15&0.75&539&0.33&1,656 \\
Maryland&5.83&0.61&4,527&0.27&9,726 \\
Massachusetts&5.10&0.54&7,975&0.24&14,954 \\
Minnesota&10.15&1.07&-774&0.47&6,843 \\
New Jersey&8.97&0.94&1,323&0.42&15,722 \\
New York&6.89&0.73&23,200&0.32&65,365 \\
Oregon&9.99&1.05&-315&0.46&3,752 \\
Pennsylvania&3.07&0.32&17,662&0.14&25,452 \\
Rhode Island&5.99&0.63&1,391&0.28&3,093 \\
Vermont&8.80&0.93&86&0.41&788 \\
Washington&0.00&0.00&11,700&0.00&13,303 \\
Average&6.75&0.71&5,810&0.31&13,422 \\