### 2023-12-12 ReplicationPackage/tables/TableB4.tex

& Forbes Population & Mean Wealth & 1982-2017 Change\\City & in 2017& in 2017 (mil) & in Forbes Population\\
\midrule
Atlanta-Sandy Springs-Gainesville, GA-AL&9&4767&2 \\
Austin-Round Rock-Marble Falls, TX&4&8025&4 \\
Birchwood&1&4200&1 \\
Bloomington, IN&1&7500&1 \\
Boston-Worcester-Manchester, MA-RI-NH&8&5988&-2 \\
Chicago-Naperville-Michigan City, IL-IN-WI&14&3443&-2 \\
Columbia, MO&2&6800&2 \\
Dallas-Fort Worth, TX&18&5950&-9 \\
Denver-Aurora-Boulder, CO&4&9900&-1 \\
Detroit-Warren-Flint, MI&3&4400&-1 \\
Fayetteville-Springdale-Rogers, AR-MO&4&20375&3 \\
Houston-Baytown-Huntsville, TX&11&4464&-10 \\
Indianapolis-Anderson-Columbus, IN&2&2700&0 \\
Jackson, WY&4&11700&4 \\
Kalamazoo-Portage, MI&2&3950&0 \\
Knoxville-Sevierville-La Follette, TN&2&3050&2 \\
Las Vegas-Paradise-Pahrump, NV&7&7471&4 \\
Los Angeles-Long Beach-Riverside, CA&31&4806&3 \\
Miami-Fort Lauderdale-Pompano Beach, FL&25&4416&13 \\
Milwaukee-Racine-Waukesha, WI&4&3800&3 \\
Minneapolis-St. Paul-St. Cloud, MN-WI&2&4100&-5 \\
Naples-Marco Island, FL&3&5033&2 \\
Nashville-Davidson--Murfreesboro--Columbia, TN&4&4275&3 \\
New York-Newark-Bridgeport, NY-NJ-CT-PA&80&6368&-9 \\
Oklahoma City-Shawnee, OK&3&7867&-2 \\
Omaha-Council Bluffs-Fremont, NE-IA&2&41100&1 \\
Philadelphia-Camden-Vineland, PA-NJ-DE-MD&5&3500&-16 \\
Phoenix-Mesa-Glendale, AZ&5&2780&5 \\
Portland-Vancouver-Hillsboro, OR-WA&2&14450&0 \\
Raleigh-Durham-Cary, NC&2&6700&2 \\
Rochester-Batavia-Seneca Falls, NY&2&2850&2 \\
San Diego-Carlsbad-San Marcos, CA&2&3750&-2 \\
San Jose-San Francisco-Oakland, CA&54&8154&37 \\
Santa Barbara-Santa Maria-Goleta, CA&2&4250&2 \\
Seattle-Tacoma-Olympia, WA&8&29775&7 \\
St. Louis-St. Charles-Farmington, MO-IL&2&5450&1 \\
Tampa-St. Petersburg-Clearwater, FL&4&2375&3 \\
Tulsa-Bartlesville, OK&2&5350&0 \\
Washington-Baltimore-Northern Virginia, DC-MD-VA-WV&10&5490&4 \\
Average&9&7470&1 \\