### 2023-12-18 template/paper.qmd
Finished fuzzy matching, began running DDDs and built summary table.

---
title: "Test"
author: "Ian Berlin"
format:
  cmc-article-pdf:
    fontsize: 12pt
    margin-left: 1.5in
    margin-right: 1.5in
    margin-bottom: 1.2in
    margin-top: 1.2in
    keep-tex: true
    fig-pos: 'tbp'
    tbl-pos: 'tbp'
    include-in-header: header.tex
bibliography: aqrd-bibliography.bib
abstract: | 
  Recent studies have shown that wealth among the 400 richest Americans — the _Forbes_ 400 — has grown increasingly concentrated among individuals who inherited their wealth.  At the same time, these individuals tend to be more stable within the Forbes 400 over time.  One potential explination for this divergence could be that individuals who belong to dynastic fortunes are better able to collude to avoid state level estate taxes. Building off of [@moretti2023], I replicate their findings that billionaires are generally sensitive to estate taxes and move between states to avoid taxation, but find that there is no difference in the rate of movement depending on whether and invidiaul belongs to a dynasty or not.  
---

# Introduction

Economic inequality is one of the biggest problems plaguing today's world. The richest 1% of people own almost 46% of the world's wealth while 55% of adults have less than \$10,000; more than half of the world owns only 1.3% of its wealth.[^1] 
One proposed mechanism for addressing growing economic inequality is inheritance taxation. Some politicians have suggested that it can raise revenue and reduce inequality of opportunity for future generations, especially when the wealthiest citizens are taxed at higher rates [@wilson2023]. **SK: This paragraph uses a footnote and a citation.**

[^1]: "Global Inequality." 2022. *Inequality.org*. <https://inequality.org/facts/global-inequality/>.

There is some disagreement about what political factors influence redistributive taxes. Some studies find that democratization leads to increased taxation while others argue that democracy and economic redistribution are only related under certain political conditions . On the other hand, some argue that democratization does not increase redistributive taxes, including, who argue that mass mobilization is a main determinant of the implementation of inheritance taxation. Using universal male suffrage and competitive elections indicators, they do not find a positive relationship between democracy and the top rate of inheritance taxation. **SK: This paragraph uses multiple citations in a group. See <https://quarto.org/docs/authoring/footnotes-and-citations.html> for more.**

# Data and Methods

For the last 40 years, _Forbes_ magazine has published an annual list of the richest individuals in America, known collectively as the _Forbes_ 400.  Through a combination of interviews and public records research, reporters attempt to construct wealth estimates for the extremely wealthy, netting their assets against debt. These data, which include names, estimate wealth, and location, are published each year, and they have been collected and cleaned by [@moretti].  This includes "extensive cleaning of the name variable in order to track individuals longitudinally."

By converting from person-year observations to state-year observations, the authors allow themselves to track the number of Forbes billionaires living in each state each year.  This structure lends itself well to a difference-in-difference (DiD) design that exploits the 2001 changes to the tax code represented by $$Y_{it} = \alpha + \beta_{1}ET_{i,t} + \beta_{2}Post + \delta ET_{i,t}\times Post + \eta_{i} + \theta_{t} + \epsilon_{it},$$ where $Y_{it}$ is the number of billionaires on the Forbes list living in state *i* in year *t*, *ET* is an indicator variable tracking whether a state has an estate tax currently enacted, *Post* is an indicator for whether the year is before or after 2001, $\eta_{i}$ controls for state fixed effects, $\theta_{t}$ controls for time fixed effects and $\epsilon_{it}$ is the error.

They construct their own variable of dynastic fortunes, largely based off of the reported source by _Forbes_, combined with additional data from [@kaplan2013] that tracks whether an individual is the first generation, second generation _etc_. to appear on the _Forbes_ list. Meanwhile, [@fernholz2023] tracks every individual longitudinally on the _Forbes_ list, grouping them together into dynasties.  One shortcoming of this measure, however, is that they treat individuals who appear in the first year of the _Forbes_ list all as seeding their own dynasty, even if they are clearly related to each other (for example, some members of the DuPont or Rockefeller families are listed as belonging to their own distinct line).

To better capture whether individuals truly belong to a dynasty or not, I have chosen to collapse these differently seeded lines, and treat all members of the same family as belonging to a dynasty, employing extensive searches online to confirm relationships between members (i.e. how many, if any, of the various Smiths on the list are part of the same dynasty?).  After coding each member with a binary dynasty variable, I used a fuzzy matching technique to append this variable to the Moretti and Wilson [-@moretti] data set.

![Time Trends for Different Measures of Democratic Participation in the United Kingdom, 1816-2000](figures/image7.png){#fig-example width="50%"}

Using data collected from V-Dem of the two new measures of democratization, joint with the data used by , I investigate the relationship between democracies that have given enough political power to their lower socioeconomic classes and redistributive taxation. I employ two main models. The first model uses a differences-in-differences approach that estimates the causal effect of democracy and wars of mass mobilization on inheritance taxation with an ordinary least squares regression represented by: $$T_{it} = \alpha + \beta_{1}D_{i,t - 1} + \beta_{2}W_{i,t - 1} + \gamma X_{i,t - 1} + \eta_{i} + \theta_{t} + \epsilon_{it},$$ where $T_{it}$ is the top inheritance tax rate for direct descendants for country *i* in year *t*.

# Results

@tbl-main shows the results of the regression. I find some evidence for the importance of considering inequities in access to power when investigating what causes increases in inheritance taxation. While inconsistent, some models find that the effect of the lagged Equal Access to Power index on top inheritance taxation is statistically significant.

[**SK: The quarto code below puts in a figure into a markdown table and calls it a table. Inserting tables are more complicated than inserting figures, so a hack is to include the image of the table as I have done here. Alternatively, you can also ask R to generate tables in a .tex format and ask quarto read it in as a tex file. See Appendix.**]

|                        |
|:-----------------------|
| ![](figures/image3.png) |

: War Mobilization, Democracy, and Inheritance Taxation, 1816-2000 {#tbl-main}

[**SK: Do not delete the two lines of quarto code that starts with `::: {#refs}`. This produces a list of references.**]

::: {#refs}
:::

# Appendix {.appendix}

```{=tex}
\begin{table}
\caption{\textbf{tex version of modelsummary output}}
\label{tab:main-tex}
\input{tables/scheve-regression.tex}
\end{table}
```

Adding a `{.appendix}` next to your section header makes it an Appendix.

Table \ref{tab:main-tex} is a table generated in LaTeX. LaTeX is a different language than R or markdown. So, we denote that we are using tex code explicitly by using the `{=tex}` tag. However, often quarto can detect LaTeX code. For example, the notation `\ref{tab:main-tex}` used in the top of this paragraph is tex code that quarto recognizes, and captures the number of the table.

The file `tables/scheve-regression.tex` can be generated in your R script. When you create the table with `modelsummary`, add `output = "gt"` inside `modelsummary()`, and saving the output of `modelsummary` into `gt::gtsave()`. For example, this uses a simple example:

![](figures/codesnippet.png){width="80%"}

Then, the code chunk starting with `{=tex}` in the quarto code will read in the tex table, and add additional environments around it. For more on these environments, see <https://www.overleaf.com/learn/latex/Environments> and the chapters in the website. ChatGPT should be quite good too at providing snippets of LaTeX code. All the teaching fellows and instructors can help with TeX.