### 2023-12-19 template/paper.qmd
Wrote literature review and additional results

---
title: "Test"
author: "Ian Berlin"
format:
  cmc-article-pdf:
    fontsize: 12pt
    margin-left: 1.5in
    margin-right: 1.5in
    margin-bottom: 1.2in
    margin-top: 1.2in
    keep-tex: true
    fig-pos: 'tbp'
    tbl-pos: 'tbp'
    include-in-header: header.tex
bibliography: aqrd-bibliography.bib
abstract: | 
  Recent studies have shown that wealth among the 400 richest Americans — the _Forbes_ 400 — has grown increasingly concentrated among individuals who inherited their wealth.  At the same time, these individuals tend to be more stable within the Forbes 400 over time.  One potential explination for this divergence could be that individuals who belong to dynastic fortunes are better able to collude to avoid state level estate taxes. Building off of @moretti_taxing_2023, I replicate their findings that billionaires are generally sensitive to estate taxes and move between states to avoid taxation, but find that there is no difference in the rate of movement depending on whether and individual belongs to a dynasty or not.  
---

# Introduction

American's billionaires are growing increasingly older (see CITE.)  Not only has billionaire wealth been growing steadily over time, wealth concentration has grown increasingly strong among the richest Americans as well [@saez_progressive_2019].  As these billionaires die, they will transfer trillions/hundreds of billions of dollars in wealth to their heirs — and many of them will do it tax free.  Plus, thanks to the stepped-up-basis, which resets the basis for capital gains at death, billions of dollars in capital gains will pass to the next generation with no ability to tax them.

Ultimately, a progressive estate tax target at the richest estates represents the best chance for fairly taxing a population that is notorious difficult to tax as a result of countless loopholes in the tax code.  Understanding how billionaires respond to an estate tax is essential as both individual states and the federal government contend with how best to design taxes ahead of a large wave of expected billionaire deaths.  In this study, I replicate evidence from @moretti_taxing_2023 that billionaires are highly sensitive to how much they pay in estate taxes in deciding where to live.  While Moretti and Wilson go on to show that this effect is much stronger for billionaires above 65 — who are much more actively engaged in estate planning — I instead look at whether or not individuals belonging to dynastic family fortunes behave differently in response to estate taxes.  I find that at least in terms of interstate migration, dynastic fortunes do not behave significantly differently than non-dynastic fortunes.

The paper proceeds as follows: Section 2 provides a literature review and motivates the focus on dynastic wealth; Section 3 presents the data and methods; Section 4 presents the results and discussion; Section 5 concludes.

# Literature Review

This paper attempts to pull together two recently emerging strands within the literature on extreme wealth.  In @korom_enduring_2017, the authors track self-made vs inherited fortunes, finding that although self-made fortunes have grown increasingly prominent within the *Forbes* 400 over time, inherited fortunes are more stable over time.  In other words, possessing an inherited fortune decreases the probability that an individual falls off the *Forbes* list in any given year, all else being equal.  At the same time, @fernholz_rising_2023 study the *Forbes* panel from 1982-2020, linking members by family dynasty.  They document a rise in wealth concentration and a decline in mobility for the wealthiest families over time.  Clearly, there is some benefit to having an inherited or dynastic fortune when trying to maintain or grow extreme wealth.  Yet the literature has done little probe potential mechanisms for this divergence.

@korom_enduring_2017 advance three potential explications: "preservation of capital through professional management of assets, tax avoidance, and the continuing control over family companies."  I focus here specifically on tax avoidance.  Although disaggregated information on tax avoidance is hard to estimate, making comparisons between family fortunes and self-made individuals difficult, we can test an extreme form of tax-avoidance: interstate migration to avoid the estate tax.  @moretti_taxing_2023 provides a valuable basis for considering estate tax avoidance among the extremely wealthy.

They first provide a valuable contribution to the literature by validating the residence locations that *Forbes* provides.  They do this in two ways: 1) by showing that the residence listed for billionaires who died during the observation period matching the location listed on their obituary 90 percent of time and 2) by showing that estate and inheritance tax revenues spike in the "correct" state following the death of a *Forbes* billionaire by an average 8.25 percent of that billionaires wealth, which is roughly half of the typical rate in ET states.  This mirrors findings from @raub_comparison_2010, which finds that billionaires report a value of wealth for tax purposes that is roughly half of the wealth Forbes reports.

@moretti_taxing_2023 then exploit changes in 2001 to the federal tax code which phased out federal credits for state estate tax liabilities, thereby introducing considerable interstate variation in total tax burden depending on location for billionaires. Using a difference-in-diffence design, they show that after 2001, billionaires on net move from ET states to non-ET states.  Furthermore, they use a triple-difference approach to consider the impact of age, showing that the effect is particularly strong for billionaires over 65.  Building off of that design, I test whether there is a higher sensitivity for dynasties over self-made individuals, thereby testing an extreme case of the tax-avoidance theory underlying the observations from @korom_enduring_2017 and @fernholz_rising_2023.

# Data and Methods

For the last 40 years, *Forbes* magazine has published an annual list of the richest individuals in America, known collectively as the *Forbes* 400. Through a combination of interviews and public records research, reporters attempt to construct wealth estimates for the extremely wealthy, netting their assets against debt. These data, which include names, estimate wealth, and location, are published each year, and they have been collected and cleaned by Moretti and Wilson [-@moretti_taxing_2023]. This includes "extensive cleaning of the name variable in order to track individuals longitudinally."

```{=tex}
\begin{table}
\caption{\textbf{Summary Statistics (Individual-Year Data)}}
\label{#tbl-one}
\input{tables/summarystats.tex}
\end{table}
```
Next, they link the *Forbes* list to state economic data, including estate tax rates, income tax rates, state GDP information, and population.  They also construct their own variable of dynastic fortunes, largely based off of the reported source by *Forbes*, combined with additional data from @kaplan_its_2013 that tracks whether an individual is the first generation, second generation *etc*. to appear on the *Forbes* list. Meanwhile, @fernholz_rising_2023 tracks every individual longitudinally on the *Forbes* list, grouping them together into dynasties. One shortcoming of this measure, however, is that they treat individuals who appear in the first year of the *Forbes* list all as seeding their own dynasty, even if they are clearly related to each other (for example, some members of the DuPont or Rockefeller families are listed as belonging to their own distinct line).  While @moretti_taxing_2023 collapse the generational information from @kaplan_its_2013 into a binary variable, it would be an fruitful area of future study to interrogate whether the results are robust to more finely grained measures of dynasty.  However, the Kaplan data only covers four years, limiting the availability of such a measure for the entire panel without building up a conceptually similar measure from scratch.

To better capture whether individuals truly belong to a dynasty or not, I have chosen to collapse these differently seeded lines, and treat all members of the same family as belonging to a dynasty, employing extensive searches online to confirm relationships between members (i.e. how many, if any, of the various Smiths on the list are part of the same dynasty?). After coding each member with a binary dynasty variable, I used a fuzzy matching technique along with manual cleaning by hand to link the data sets and append Fernholz and Hagler's dynasty codings to the Moretti and Wilson [-@moretti_taxing_2023] data set.  @tbl-one shows summary statistics for 12,694 individual-year observations spanning from 1983-2017.  1984 has been omitted because it is missing from the Fernholz and Hagler data, and the Moretti and Wilson data excludes 2002 because *Forbes* did not publish location data that year.  @fig-panel shows a summary of treatment — whether or not an estate tax is in place — for each state over the entire observation period.

![Summary View of Estate Tax Treatment by State, 1982-2017](figures/panelview.png){#fig-panel width="87%"}

By converting from person-year observations to state-year observations, the authors allow themselves to track the number of Forbes billionaires living in each state each year. Appendix Table A1 replicates Table 1 from Moretti and Wilson, and shows the structure of the state-year data, including per-state summaries of Forbes population and change over time.  This structure lends itself well to a difference-in-difference (DiD) design that exploits the 2001 changes to the tax code represented by $$Y_{it} = \alpha + \beta_{1}ET_{i,t} + \beta_{2}Post_{i,t} + \delta(ET_{i,t}\times Post_{i,t}) + \eta_{i} + \theta_{t} + \epsilon_{it},$$ where $Y_{it}$ is the number of billionaires on the Forbes list living in state *i* in year *t*, *ET* is an indicator variable tracking whether a state has an estate tax currently enacted, *Post* is an indicator for whether the year is before or after 2001, $\eta_{i}$ is state fixed effects, $\theta_{t}$ is time fixed effects and $\epsilon_{it}$ is the error.

Whereas Moretti and Wilson proceed to a triple-difference design that shows that individuals over 65 are more likely to move to avoid taxes then those under 65, I use a similar design to test whether members of dynastic fortunes are more likely to move to avoid taxes than those who are not.  This triple difference is represented by $$Y_{it} = \alpha + \beta_{1}ET_{i,t} + \beta_{2}Post_{i,t} + \beta_{3}D_{i,t} + \beta_{4} (ET_{i,t} \times Post_{i,t}) + \beta_{5} (ET_{i,t} \times D_{i,t}) +$$ $$\beta_{6} (Post_{i,t} \times D_{i,t}) + \delta (ET_{i,t}\times Post_{i,t} \times D_{i,t}) + \eta_{i} + \theta_{t} + \epsilon_{it},$$

where *D* is added to the previous equation (and interacted with each term) and represents the count of dynasty members living in state *i* during time *t*.

# Results

Following @moretti_taxing_2023, I "use the elimination of the federal credit for state estate taxes in 2001 for assessing the locational sensitivity of the ultra-rich to estate taxes."  In particular, while billionaires generally paid the same in taxes regardless of location before 2001, their tax burdens varied substantially by state post 2001.  @fig-five shows the share of billionaires living in a state that had an estate tax in 2001 each year.  The horizontal line shows the passage of the EGTRRA with the shaded region showing the phase in period.  The horizontal lines correspond to the pre-2001 and post-2001 mean, showing a drop from about 21 percent on average to roughly 17 percent. @fig-five replicates Figure 5 from @moretti_taxing_2023.

![Share of Forbes 400 Living in a 2001 Estate Tax State, 1982-2017](figures/figure5.png){#fig-five}
To test this relationship more rigorously, I replicate Moretti and Wilson's [-@moretti_taxing_2023] difference-in-difference design that compares the change in the number of billionaires  prior to the 2001 credit elimination and after the elimination in estate tax states and non–estate tax states.  Note that while the estimates match the original paper precisely, Moretti and Wilson use Driscoll-Kraay (with ten-year bandwidth) standard errors, whereas I use simple clustered standard errors, thus explaining the difference.  Column numbers correspond to the original column in the paper.

|                         |
|:------------------------|
| ![](tables/reg1.png) | 
: Partial replication of @moretti_taxing_2023 Table 2: Difference-in-Difference Dependent Variable: Population of Forbes 400 {#tbl-2}

Across all columns, the interaction term is negative and significant, showing that "the average estate tax state saw a drop of billionaires after 2001 compared with the average non–estate tax state" [@moretti_taxing_2023].  In column 1, which presents the simple model, the coeffcient is -2.454, which corresponds to a loss of 2.45 billionaires in estate tax states after 2001 compared to non-estate tax states.  However, states often change numerous tax laws at the same time, so migration in response to these other changes may appear to be migration in response to the estate tax.  Column 2 controls for the top marginal personal income tax (PIT) rate by state and year, and the findings are robust to this addition.  Controlling for PIT, states with estate taxes lose 2.68 billionaires after 2001 compared to non-estate tax states.

Another potential concern is that "the amenity value of non-ET states relative to ET states has increased over time for the very wealthy, leading to an increase over time in the share of the *Forbes* 400 living in non-ET states for reasons unrelated to the estate tax."  Therefore, column 3 controls for the share of high earners — individuals between the 90th and 97th percentile — in a state who would similarly benefit from amenities but do not pay the estate tax.  The negative coefficient on the interaction term (-2.32) is largely unchanged, showing that the decrease in population in ET states is not driven by improved amenities in non-ET states.  Column 5 uses per capita *Forbes* billionaires per state as the dependent variable, finding a coefficient of -6.21.  Column 8 shows that the results are robust to dropping 2002-2004 from the dataset, which is the period during which federal tax credits were being phased out.  @Moretti_taxing_2023 also show that their estimates are robust to replacing their estate tax dummy with specific tax rates, as well as subsetting by only the top 100, 200, and 300 wealthiest individuals in order to remove individuals who are more likely to drop off the *Forbes* list.

I then move to the triple-difference design.  @tbl-trip presents the results. None of the coefficients on the three-way interaction term are significant, suggesting that members of dynastic fortunes do not move to avoid the estate tax significantly more than non-dynastic individuals.  Columns 1-4 match the same first four columns of @tbl-2, controlling for PIT rate (and its interactions), high earner share, and number of billionaires per capita.  The next column uses the alternative measure of dynasty from @moretti_taxing_2023, while the following two columns show the results from combinations of the two, first coding an individual as dynsatic if either measure codes them as such and second coding and individual as dynastic if both measures code them as such.  The three-way interaction is non-significant for all three.  Finally, the last column shows that dropping the phase in period of 2002-2004 does not alter the results.

|                         |
|:------------------------|
| ![](tables/reg2.png){#fig-reg width="90%"}| 
: Triple-Difference Dependent Variable: Population of Forbes 400 {#tbl-trip}

# Conclusion

::: {#refs}
:::

# Appendix {.appendix}

|                         |
|:------------------------|
| ![](figures/table1.png) |

: Replication of @moretti_taxing_2023 Figure 1: "Forbes 400 by State, 2017" {#tbl-main}

![Replication of @moretti_taxing_2023 Figure 6: "Probability of Living in Estate Tax State by Age](figures/figure_6.png){#fig-6}

Adding a `{.appendix}` next to your section header makes it an Appendix.

Table \ref{tab:main-tex} is a table generated in LaTeX. LaTeX is a different language than R or markdown. So, we denote that we are using tex code explicitly by using the `{=tex}` tag. However, often quarto can detect LaTeX code. For example, the notation `\ref{tab:main-tex}` used in the top of this paragraph is tex code that quarto recognizes, and captures the number of the table.

The file `tables/scheve-regression.tex` can be generated in your R script. When you create the table with `modelsummary`, add `output = "gt"` inside `modelsummary()`, and saving the output of `modelsummary` into `gt::gtsave()`. For example, this uses a simple example:

Then, the code chunk starting with `{=tex}` in the quarto code will read in the tex table, and add additional environments around it. For more on these environments, see <https://www.overleaf.com/learn/latex/Environments> and the chapters in the website. ChatGPT should be quite good too at providing snippets of LaTeX code. All the teaching fellows and instructors can help with TeX.