### 2023-12-20 template/paper.qmd
final update

---
title: "Preserving Fortunes: Dynastic Wealth, Taxation, and Interstate Mobility"
author: "Ian Berlin"
format:
  cmc-article-pdf:
    fontsize: 12pt
    margin-left: 1.5in
    margin-right: 1.5in
    margin-bottom: 1.2in
    margin-top: 1.2in
    keep-tex: true
    fig-pos: 'tbp'
    tbl-pos: 'tbp'
    include-in-header: header.tex
bibliography: aqrd-bibliography.bib
abstract: | 
  Recent studies have shown that wealth among the 400 richest Americans — the *Forbes* 400 — has grown increasingly concentrated over time.  At the same time, individuals who inherited their wealth tend to be more stable within the *Forbes* 400 over time.  One potential explanation for this divergence could be that individuals who belong to dynastic fortunes are better at colluding to avoid state level estate taxes by positioning where they live. Building off of @moretti_taxing_2023, I replicate their findings that billionaires are generally sensitive to estate taxes and move between states to avoid taxation, but find that there is no difference in the rate of movement depending on whether an individual belongs to a dynasty or not.  
---

# Introduction

American's billionaires are growing increasingly older.[^1]  Meanwhile, fortunes have grown increasingly concentrated at the very top of the wealth distribution  [@saez_progressive_2019].  As these billionaires die, they will transfer trillions of dollars in wealth to their heirs — and, thanks to the stepped-up-basis, which resets the basis for capital gains at death, billions of dollars in generated wealth will pass to the next generation with no ability to tax them.

[^1]: See: https://www.cnbc.com/2023/05/31/billionaire-median-age-study.html

In theory, a progressive estate tax targeted at the richest estates represents the best chance for fairly taxing a population that is notorious evasive, employing countless loopholes to avoid paying their share.  Understanding how billionaires respond to estate taxes is essential as both individual states and the federal government consider how best to tax these massive intergenerational transfers.  In this study, I replicate evidence from @moretti_taxing_2023 that billionaires are highly sensitive to how much they pay in estate taxes in deciding where to live.  While Moretti and Wilson go on to show that this effect is stronger for billionaires above 65 — who are much more actively engaged in estate planning — I instead look at whether or not individuals belonging to dynastic family fortunes behave differently in response to estate taxes.  I find that dynastic fortunes do not behave significantly differently than non-dynastic fortunes, at least in terms of interstate migration.

The paper proceeds as follows: Section 2 provides a literature review and motivates the focus on dynastic wealth; Section 3 presents the data and methods; Section 4 presents the results and discussion; Section 5 concludes.

# Literature Review

This paper attempts to pull together two recently emerging strands within the literature on extreme wealth.  In @korom_enduring_2017, the authors track self-made vs. inherited fortunes, finding that, although self-made fortunes have grown increasingly prominent within the *Forbes* 400 over time, inherited fortunes are more stable over time.  In other words, possessing an inherited fortune decreases the probability that an individual falls off the *Forbes* list in any given year, all else being equal.  At the same time, @fernholz_rising_2023 study the *Forbes* panel from 1982-2020, linking members by family dynasty.  They document a rise in wealth concentration and a decline in mobility for the wealthiest families over time.  These results show that there is some benefit to having an inherited or dynastic fortune when trying to maintain or grow extreme wealth.  Yet the literature has done little to probe potential mechanisms for this divergence.

@korom_enduring_2017 advance three potential explanations: the "preservation of capital through professional management of assets, tax avoidance, and the continuing control over family companies."  I focus here specifically on tax avoidance.  Although specific information on individual tax avoidance is hard to estimate, making direct comparison between dynastic and non-dynastic fortunes difficult, we can test an extreme form of tax-avoidance: interstate migration to avoid the state-level estate tax entirely.  @moretti_taxing_2023 provide a valuable basis for considering estate tax avoidance among the extremely wealthy.

They first provide a critical contribution to the literature by validating the residence information that *Forbes* provides.  They do this in two ways: 1) by showing that the residence listed for billionaires who died during the observation period matched the location listed on their obituary 90 percent of time and 2) by showing that estate and inheritance tax revenues spike in the "correct" state following the death of a *Forbes* billionaire by an average of 8.25 percent of that billionaire's wealth.  That is roughly half of the typical estate tax rate in ET states, mirroring findings from @raub_comparison_2010, who find that billionaires report a wealth value for tax purposes that is roughly half of the wealth that *Forbes* reports.

Having demonstrated the validity of using *Forbes*'s location information, @moretti_taxing_2023 then exploit changes to the federal tax code under the 2001 Bush tax cuts.  These changes  phased out federal credits for state estate tax liabilities, thereby introducing considerable interstate variation in total tax burden depending which state a billionaire lived in. Using a difference-in-difference design, they show that, after 2001, billionaires on net move from ET states to non-ET states.  Furthermore, they use a triple-difference approach to consider the impact of age, showing that the effect is particularly strong for billionaires over 65 (see @fig-6 in the appendix for a replication of these findings visually).  Building off of that design, I probe whether there is a higher sensitivity for dynasties over self-made individuals, thereby testing an extreme case of the tax-avoidance theory underlying the observations from @korom_enduring_2017 and @fernholz_rising_2023.

# Data and Methods

For the last 40 years, *Forbes* magazine has published an annual list of the richest individuals in America, known collectively as the *Forbes* 400. Through a combination of interviews and public records research, reporters attempt to construct wealth estimates for the extremely wealthy, subtracting their debts from their assets. These data, which include names, estimated wealth, and location, are published each year, and they have been collected and cleaned by @moretti_taxing_2023. This includes "extensive cleaning of the name variable in order to track individuals longitudinally."

Next, they link the *Forbes* list to state economic data, including estate tax rates, income tax rates, state GDP information, and population.  They also construct their own variable of dynastic fortunes, based on whether *Forbes* lists the source as "inherited" or not.  Meanwhile, @fernholz_rising_2023 track every individual longitudinally on the *Forbes* list, manually grouping them together into dynasties through careful internet investigation. One shortcoming of this measure, however, is that they treat individuals who appear in the first year of the *Forbes* list  as each seeding their own dynasty, even if they are clearly related to each other (for example, several members of the DuPont or Rockefeller families belong to their own "dynasty" because they were already wealthy in the first year of the *Forbes* list and never married or had children).  Therefore, I conduct additional research to confirm whether such individuals in fact belong to a dynasty and code them as such, even if they appear as an individual fortune with in the @fernholz_rising_2023 dataset.  Furthermore, @kaplan_its_2013 introduce a generational measure of dynasty, which codes individuals as first, second, third *etc.* generation.  While a more finely grained measure of dynastic wealth would be useful, the Kaplan data only covers four years of the panel, limiting its utility within the scope of this paper.  A worthwhile future area of investigation could involve expanding their measure or something conceptually similar to all years within the *Forbes* panel.

|                         |
|:------------------------|
| ![](tables/summarystats.png) | 
: Summary Statistics {#tbl-summary}

After coding each member with a binary dynasty variable based on the @fernholz_rising_2023 data, I used a fuzzy matching technique using the `jaccard_join` command in the `zoomerjoin` package, followed by extensive manual cleaning by hand to append the dynasty codings to the @moretti_taxing_2023 data set.  Given that the @fernholz_rising_2023 metric was constructed via independent research, whereas @moretti_taxing_2023 simply adopt *Forbes*'s broad "source" information, it will be the preferred measure for my analysis. @tbl-summary shows summary statistics for 12,694 individual-year observations spanning from 1983-2017.  1984 has been omitted because it is missing from the @fernholz_rising_2023 data, and the @moretti_taxing_2023 data excludes 2002 because *Forbes* did not publish location data that year.  Additionally, some individuals have been excluded from the @moretti_taxing_2023 data in certain years because *Forbes* failed to report a location that year.  @fig-panel shows a summary of treatment — whether or not an estate tax is in place — for each state over the entire observation period.

By converting from person-year observations to state-year observations, the authors allow themselves to track the number of *Forbes* billionaires living in each state each year. @tbl-main (see appendix) replicates Table 1 from @moretti_taxing_2023, and shows the structure of the state-year data, including per-state summaries of the *Forbes* population in 2017 and change over time.  This structure lends itself well to a difference-in-difference design that exploits the 2001 changes to the tax code, represented by $$Y_{it} = \alpha + \beta_{1}ET_{it} + \beta_{2}Post_{it} + \delta(ET_{it}\times Post_{it}) + \eta_{i} + \theta_{t} + \epsilon_{it},$$ where $Y_{it}$ is the number of billionaires on the *Forbes* list living in state *i* in year *t*, *ET* is an indicator variable tracking whether a state has an estate tax enacted, *Post* is an indicator for whether the year is before or after 2001, $\eta_{i}$ is state fixed effects, $\theta_{t}$ is time fixed effects and $\epsilon_{it}$ is the error.

Whereas @moretti_taxing_2023 proceed to a triple-difference design that further differences by age (above or below 65), I employ a similar design to instead introduce the dynasty measures discussed above as the additional difference.  This triple difference is represented by 
$$Y_{it} = \alpha + \beta_{1}ET_{it} + \beta_{2}Post_{it} + \beta_{3}D_{it} + \beta_{4} (ET_{it} \times Post_{it}) + \beta_{5} (ET_{it} \times D_{it}) +$$ $$\beta_{6} (Post_{it} \times D_{it}) + \delta (ET_{it}\times Post_{it} \times D_{it}) + \eta_{i} + \theta_{t} + \epsilon_{it},$$

where *D* is added to the previous equation (and interacted with each term) and represents whether an individual belongs to a dynasty or not.

![Summary View of Estate Tax Treatment by State, 1982-2017](figures/panelview.png){#fig-panel width="85%"}

Whereas the typical difference-in-difference design requires the parallel trends assumption to be met, the triple-difference estimator does not require two parallel trends.  This is because the "difference between two biased difference-in-differences estimators will be unbiased as long
as the bias is the same in both estimators" [@olden_triple_2022].  In the case of my model, even if there are time *and* state variant confounding variables affecting interstate mobility, they will be captured by the triple-difference as long as they affect dynastic and non-dynastic individuals similarly.

# Results

Following @moretti_taxing_2023, I "use the elimination of the federal credit for state estate taxes in 2001 for assessing the locational sensitivity of the ultra-rich to estate taxes."  In particular, while billionaires generally paid the same in taxes regardless of location before 2001, their tax burdens varied substantially by state post-2001.  @fig-per (replicating Figure 5 from @moretti_taxing_2023) shows the share of billionaires living in a state that had an estate tax in 2001 over time.  The horizontal line shows the passage of the Bush tax cuts, with the shaded region showing the phase out period for the ET credits.  The horizontal lines correspond to the pre-2001 and post-2001 means, showing a drop from about 21 percent on average to roughly 17 percent.

![Replication of Figure 5: "Share of *Forbes* 400 Living in a 2001 Estate Tax State"](figures/figure5.png){#fig-per width="100%"}

To test this relationship more rigorously, I replicate @moretti_taxing_2023 difference-in-difference design that compares the change in the number of billionaires  prior to the 2001 credit elimination and after the elimination in estate tax states and non–estate tax states.  Note that while the estimates match the original paper precisely, Moretti and Wilson use Driscoll-Kraay (with ten-year bandwidth) standard errors, whereas I use simple clustered standard errors, thus explaining the difference.  Column numbers correspond to the original column in the paper as some columns are omitted because they fall outside the scope of this analysis.

|                         |
|:------------------------|
| ![](tables/reg1.png) | 
: Partial replication of @moretti_taxing_2023 Table 2: "Difference-in-Difference Dependent Variable: Population of *Forbes* 400" {#tbl-2}

Across all columns, the interaction term is negative and significant, showing that "the average estate tax state saw a drop of billionaires after 2001 compared with the average non–estate tax state" [@moretti_taxing_2023].  In column 1, which presents the simple model, the coefficient is -2.454, corresponding to a loss of 2.45 billionaires in estate tax states after 2001 compared to non-estate tax states.  However, states often change numerous tax laws at the same time, so migration in response to these other changes may mistakenly appear to be migration in response to the estate tax.  Column 2 controls for the top marginal personal income tax (PIT) rate by state and year, and the findings are robust to this addition.  Controlling for PIT, states with estate taxes lose an average of 2.68 billionaires after 2001 compared to non-estate tax states.

Another potential concern is that "the amenity value of non-ET states relative to ET states has increased over time for the very wealthy, leading to an increase over time in the share of the *Forbes* 400 living in non-ET states for reasons unrelated to the estate tax" [@moretti_taxing_2023].

Therefore, column 3 controls for the share of high earners in a state — individuals between the 90th and 97th percentile —who would similarly benefit from amenities but do not pay the estate tax.  The negative coefficient on the interaction term (-2.32) is largely unchanged, showing that the decrease in population in ET states is not driven by improved amenities in non-ET states.  Column 5 uses per capita *Forbes* billionaires per state as the dependent variable, finding a coefficient of -0.62.  Column 8 shows that the results are robust to dropping 2002-2004 from the dataset, which is the period during which the federal tax credits were being phased out.  @moretti_taxing_2023 also show that their estimates are robust to replacing their estate tax dummy with specific tax rates, as well as subsetting by only the top 100, 200, and 300 wealthiest individuals in order to remove individuals who are more likely to drop off the *Forbes* list.

I then move to the triple-difference design.  @tbl-trip presents the results. None of the coefficients on the three-way interaction term are significant, suggesting that members of dynastic fortunes do not move to avoid the estate tax significantly more than non-dynastic individuals.  Columns 1-4 match the same first four columns of @tbl-2, controlling for PIT rate (and its interactions) and high earner share, then using per capita billionaires as the dependent variable.  The next column uses the alternative measure of dynasty from @moretti_taxing_2023, while the following two columns show the results if combinations of the two are used instead. The first codes an individual as dynastic if either measure codes them as such, and the second codes an individual as dynastic only if both measures code them as such.  The three-way interaction is non-significant for all three. Therefore, results are robust to using different theoretical definitions of dynasty.  Finally, the last column shows that dropping the phase in period of 2002-2004 does not alter the results.

|                         |
|:------------------------|
| ![](tables/reg2.png){#fig-reg width="90%"}| 
: Triple-Difference Dependent Variable: Population of *Forbes* 400 {#tbl-trip}

# Conclusion

This paper makes an important foray into testing plausible mechanisms to explain the relative stability of dynastic fortunes over time compared to self-made individuals.  Although my results suggest that location based gamesmanship in order to evade the state-level estate tax is not a plausible explanation, future studies could test less extreme cases of tax avoidance along with non-taxation based explanations, such as better asset management.  While accurate data is likely less readily available for such approaches, efforts to better understand inter-generational wealth transfers will help inform policy makers as they design economic policies aimed at taxing the extremely wealthy.

::: {#refs}
:::

# Appendix {.appendix}

|                         |
|:------------------------|
| ![](figures/table1.png) |

: Replication of @moretti_taxing_2023 Figure 1: "*Forbes* 400 by State, 2017"{#tbl-main}

![Replication of @moretti_taxing_2023 Figure 6: "Probability of Living in Estate Tax State by Age"](figures/figure_6.png){#fig-6}